---
title: "Model Specification and Sensitivity Analysis"
author: "Siard van den Bosch"
date: last-modified
---

```{r}
#| label: setup
#| echo: FALSE
#| include: FALSE

rm(list = ls()) # remove all objects from env
invisible(gc(full = TRUE)) # free up memory

options(digits = 3, scipen = 100, knitr.kable.NA = "—") # set global options
knitr::opts_knit$set( # configure knitr options
  progress = FALSE,
  verbose = FALSE,
  cache = FALSE,
  knitr.table.format = "html",
  escape = FALSE,
  digits = 3,
  row.names = FALSE
)
```

```{r}
#| label: load-project-resource
#| echo: FALSE
functions <- c("pipes", "format_p_value", "named_list", "load_most_recent_rds", "describe", "numeric_data_frame", "create_formula", "save_rds", "move_column_after", "compare_models", "rownames_to_column", "standardized_residuals", "remove_columns_by_name", "beautify_names", "calculate_performance", "model_info", "compare_nested_models", "structure_models", "essential", "bootstrap", "model_summary", "format_coef", "highlight_function", "second_order_interactions", "johnson_neyman_plot", "jn_plot", "extract_term_ranges", "calculate_p_value_from_ci", "is_continuous", "footnote", "as_kable", "add_style")

purrr::walk(functions, ~ source(file.path("functions", paste0(.x, ".R"))))
unadjusted <- haven::read_sav("data/imputed_data.sav")
adjusted <- NULL; #load_most_recent_rds("adjusted_data", "data")

unadjusted_fits <- NULL #load_most_recent_rds("unadjusted_fits_", dir = "data")
adjusted_fits <- NULL #load_most_recent_rds("adjusted_fits_", dir = "data")

# SET GLOBAL N BOOTSTRAP REPLICATES
N_BOOTSTRAPS <- 1000
```


```{r eval=FALSE, message=FALSE, warning=FALSE}
# purrr::iwalk(functions, ~ highlight_function(.x))
```

## Model Specification

Model 1 examines the association between SEP and diet quality (DHD index)
```{r}
#| label: model-1
model_1 <- dhd_index ~ education_self + occupation_self
```

Model 2 investigates how SEP, household income, and financial difficulties relate to diet quality
```{r}
#| label: model-2
# Removed terms: No terms were removed
# Added terms: income_hh_eq, financial_difficulties
model_2 <- dhd_index ~ education_self + occupation_self +
  income_hh_eq + financial_difficulties
```

Model 3 examines gender differences in the relationship between individual SEP and diet quality
```{r}
#| label: model-3
# Removed terms: No terms were removed
# Added terms: woman, education_self:woman, occupation_self:woman
model_3 <- dhd_index ~ education_self + occupation_self + woman +
  education_self:woman + occupation_self:woman +
  income_hh_eq + financial_difficulties
```

Model 4 explores the relationship between SEP and diet quality, accounting for parental SEP
```{r}
#| label: model-4
# Removed terms: No terms were removed
# Added terms: education_father, occupation_father, education_mother, occupation_mother
model_4 <- dhd_index ~ education_self + occupation_self +
  education_father + occupation_father + education_mother +
  occupation_mother + income_hh_eq + financial_difficulties
```

Model 5 includes interactions between the individual's SEP and parental SEP
```{r}
#| label: model-5
# Removed terms: No terms were removed
# Added terms: education_self:education_father, education_self:education_mother,
# occupation_self:occupation_father, occupation_self:occupation_mother
model_5 <- dhd_index ~ education_self + occupation_self +
  education_father + occupation_father + education_mother +
  occupation_mother +
  education_self:education_father + education_self:education_mother +
  occupation_self:occupation_father + occupation_self:occupation_mother +
  income_hh_eq + financial_difficulties
```

Model 6 examines if gender moderates the relationship between SEP and diet quality

```{r}
#| label: model-6
# Removed terms: No terms were removed
# Added terms: education_self:woman, occupation_self:woman,
# education_father:woman, occupation_father:woman,
# education_mother:woman, occupation_mother:woman
model_6 <- dhd_index ~ education_self + occupation_self + woman +
  education_father + occupation_father + education_mother +
  occupation_mother +
  education_self:woman + occupation_self:woman +
  education_father:woman + occupation_father:woman +
  education_mother:woman + occupation_mother:woman +
  income_hh_eq + financial_difficulties
```



Model 7 includes interactions between the individual's SEP and partner's SEP

```{r}
#| label: model-7
# Removed terms: No terms were removed
# Added terms: education_self:education_partner, occupation_self:occupation_partner
model_7 <- dhd_index ~ education_self + occupation_self +
  education_partner + occupation_partner +
  education_self:education_partner + occupation_self:occupation_partner +
  income_hh_eq + financial_difficulties
```

Model 8 includes interactions between the individual's SEP, partner's SEP, and gender

```{r}
#| label: model-8
# Removed terms: No terms were removed
# Added terms: education_self:woman, occupation_self:woman,
# education_partner:woman, occupation_partner:woman
model_8 <- dhd_index ~ education_self + occupation_self + woman +
  education_partner + occupation_partner +
  education_self:education_partner + occupation_self:occupation_partner +
  education_self:woman + occupation_self:woman +
  education_partner:woman + occupation_partner:woman +
  income_hh_eq + financial_difficulties
```

Model 9 is a multivariate model including individual characteristics and justified interactions

```{r}
#| label: model-9
# Removed terms: No terms were removed
# Added terms: age, smoking_status, dhd_kcal, hh_size, children_yn,
# education_self:education_father, education_self:education_mother,
# occupation_self:occupation_father, occupation_self:occupation_mother,
# education_self:education_partner, occupation_self:occupation_partner,
# education_self:woman, occupation_self:woman
model_9 <- dhd_index ~ education_self + occupation_self + woman +
  education_father + occupation_father + education_mother +
  occupation_mother + education_partner + occupation_partner +
  age + smoking_status + dhd_kcal + hh_size + children_yn +
  education_self:education_father + education_self:education_mother +
  occupation_self:occupation_father + occupation_self:occupation_mother +
  education_self:education_partner + occupation_self:occupation_partner +
  education_self:woman + occupation_self:woman +
  education_partner:woman + occupation_partner:woman +
  income_hh_eq + financial_difficulties
```

Certainly! Below are 10 alternative models for each of the 9 initial models, designed to align with your research questions and goals. These models offer different combinations of variables and interactions to help you determine the best model specifications. Each model includes the necessary R code.

```{r}
#| label: alternative-models

# Model 1: Association Between Individual SEP and Diet Quality

# Model 1a: Add Age as a Covariate


model_1a <- dhd_index ~ education_self + occupation_self + age

# Model 1b: Include Gender


model_1b <- dhd_index ~ education_self + occupation_self + woman

# Model 1c: Add Smoking Status


model_1c <- dhd_index ~ education_self + occupation_self + smoking_status

# Model 1d: Interaction Between Education and Occupation


model_1d <- dhd_index ~ education_self * occupation_self

# Model 1e: Include Household Size


model_1e <- dhd_index ~ education_self + occupation_self + hh_size

# Model 1f: Add Children in Household


model_1f <- dhd_index ~ education_self + occupation_self + children_yn

# Model 1g: Include Caloric Intake (dhd_kcal)


model_1g <- dhd_index ~ education_self + occupation_self + dhd_kcal

# Model 1h: Add Financial Difficulties


model_1h <- dhd_index ~ education_self + occupation_self + financial_difficulties

# Model 1i: Interaction Between Education and Gender


model_1i <- dhd_index ~ education_self * woman + occupation_self

# Model 1j: Include Both Age and Smoking Status


model_1j <- dhd_index ~ education_self + occupation_self + age + smoking_status

# Model 2: Adding Household Income and Financial Difficulties

# Model 2a: Interaction Between Education and Income


model_2a <- dhd_index ~ education_self + occupation_self + income_hh_eq +
  financial_difficulties + education_self:income_hh_eq

# Model 2b: Interaction Between Occupation and Financial Difficulties


model_2b <- dhd_index ~ education_self + occupation_self + income_hh_eq +
  financial_difficulties + occupation_self:financial_difficulties

# Model 2c: Include Age and Gender


model_2c <- dhd_index ~ education_self + occupation_self + income_hh_eq +
  financial_difficulties + age + woman

# Model 2d: Add Smoking Status


model_2d <- dhd_index ~ education_self + occupation_self + income_hh_eq +
  financial_difficulties + smoking_status

# Model 2e: Interaction Between Income and Financial Difficulties


model_2e <- dhd_index ~ education_self + occupation_self + income_hh_eq *
  financial_difficulties

# Model 2f: Include Household Size and Children


model_2f <- dhd_index ~ education_self + occupation_self + income_hh_eq +
  financial_difficulties + hh_size + children_yn

# Model 2g: Add Home Ownership Status


model_2g <- dhd_index ~ education_self + occupation_self + income_hh_eq +
  financial_difficulties + home_owner

# Model 2h: Interaction Between Education and Financial Difficulties


model_2h <- dhd_index ~ education_self * financial_difficulties +
  occupation_self + income_hh_eq

# Model 2i: Include Diet-Related Variables


model_2i <- dhd_index ~ education_self + occupation_self + income_hh_eq +
  financial_difficulties + dhd_kcal

# Model 2j: Interaction Between Occupation and Income


model_2j <- dhd_index ~ education_self + occupation_self * income_hh_eq +
  financial_difficulties

# Model 3: Gender Differences in the Relationship Between SEP and Diet Quality

# Model 3a: Add Interaction Between Income and Gender


model_3a <- dhd_index ~ education_self + occupation_self + woman +
  income_hh_eq + financial_difficulties + income_hh_eq:woman

# Model 3b: Include Interaction Between Financial Difficulties and Gender


model_3b <- dhd_index ~ education_self + occupation_self + woman +
  income_hh_eq + financial_difficulties + financial_difficulties:woman

# Model 3c: Interaction Between Education, Occupation, and Gender


model_3c <- dhd_index ~ education_self * occupation_self * woman +
  income_hh_eq + financial_difficulties

# Model 3d: Add Age and Smoking Status


model_3d <- dhd_index ~ education_self + occupation_self + woman +
  income_hh_eq + financial_difficulties + age + smoking_status

# Model 3e: Interaction Between Education and Occupation


model_3e <- dhd_index ~ education_self * occupation_self + woman +
  income_hh_eq + financial_difficulties

# Model 3f: Include Household Size and Children


model_3f <- dhd_index ~ education_self + occupation_self + woman +
  income_hh_eq + financial_difficulties + hh_size + children_yn

# Model 3g: Interaction Between Education, Gender, and Financial Difficulties


model_3g <- dhd_index ~ education_self * woman * financial_difficulties +
  occupation_self + income_hh_eq

# Model 3h: Add Home Ownership Status


model_3h <- dhd_index ~ education_self + occupation_self + woman +
  income_hh_eq + financial_difficulties + home_owner

# Model 3i: Interaction Between Occupation and Gender Only


model_3i <- dhd_index ~ education_self + occupation_self * woman +
  income_hh_eq + financial_difficulties

# Model 3j: Include Diet-Related Variables


model_3j <- dhd_index ~ education_self + occupation_self + woman +
  income_hh_eq + financial_difficulties + dhd_kcal

# Model 4: Including Parental Socioeconomic Position

# Model 4a: Add Interaction Between Parental Education


model_4a <- dhd_index ~ education_self + occupation_self +
  education_father * education_mother + income_hh_eq + financial_difficulties

# Model 4b: Interaction Between Parental Occupation


model_4b <- dhd_index ~ education_self + occupation_self +
  occupation_father * occupation_mother + income_hh_eq + financial_difficulties

# Model 4c: Include Gender


model_4c <- dhd_index ~ education_self + occupation_self + woman +
  education_father + occupation_father + education_mother +
  occupation_mother + income_hh_eq + financial_difficulties

# Model 4d: Interaction Between Parental Education and Gender


model_4d <- dhd_index ~ education_self + occupation_self +
  education_father + education_mother + woman +
  education_father:woman + education_mother:woman +
  income_hh_eq + financial_difficulties

# Model 4e: Add Age and Smoking Status


model_4e <- dhd_index ~ education_self + occupation_self +
  education_father + occupation_father + education_mother +
  occupation_mother + income_hh_eq + financial_difficulties +
  age + smoking_status

# Model 4f: Interaction Between Individual and Parental Education


model_4f <- dhd_index ~ education_self * education_father +
  education_self * education_mother + occupation_self +
  occupation_father + occupation_mother + income_hh_eq + financial_difficulties

# Model 4g: Include Household Size and Children


model_4g <- dhd_index ~ education_self + occupation_self +
  education_father + occupation_father + education_mother +
  occupation_mother + income_hh_eq + financial_difficulties +
  hh_size + children_yn

# Model 4h: Interaction Between Parental Occupation and Gender


model_4h <- dhd_index ~ education_self + occupation_self +
  occupation_father + occupation_mother + woman +
  occupation_father:woman + occupation_mother:woman +
  income_hh_eq + financial_difficulties

# Model 4i: Add Financial Difficulties Interaction


model_4i <- dhd_index ~ education_self + occupation_self +
  education_father + occupation_father + education_mother +
  occupation_mother + income_hh_eq + financial_difficulties +
  financial_difficulties:education_father +
  financial_difficulties:education_mother

# Model 4j: Include Home Ownership Status


model_4j <- dhd_index ~ education_self + occupation_self +
  education_father + occupation_father + education_mother +
  occupation_mother + income_hh_eq + financial_difficulties +
  home_owner

# Model 5: Interactions Between Individual and Parental SEP

# Model 5a: Add Interaction Between Occupation Self and Education Father


model_5a <- dhd_index ~ education_self + occupation_self +
  education_father + occupation_father + education_mother +
  occupation_mother + occupation_self:education_father +
  income_hh_eq + financial_difficulties

# Model 5b: Interaction Between Education Self and Occupation Father


model_5b <- dhd_index ~ education_self + occupation_self +
  education_father + occupation_father + education_mother +
  occupation_mother + education_self:occupation_father +
  income_hh_eq + financial_difficulties

# Model 5c: Include Gender and Its Interactions


model_5c <- dhd_index ~ education_self + occupation_self +
  education_father + occupation_father + education_mother +
  occupation_mother + woman +
  education_self:education_father + education_self:education_mother +
  occupation_self:occupation_father + occupation_self:occupation_mother +
  woman:education_self + woman:occupation_self +
  income_hh_eq + financial_difficulties

# Model 5d: Add Age and Smoking Status


model_5d <- dhd_index ~ education_self + occupation_self +
  education_father + occupation_father + education_mother +
  occupation_mother + education_self:education_father +
  education_self:education_mother + occupation_self:occupation_father +
  occupation_self:occupation_mother + income_hh_eq + financial_difficulties +
  age + smoking_status

# Model 5e: Include Household Size and Children


model_5e <- dhd_index ~ education_self + occupation_self +
  education_father + occupation_father + education_mother +
  occupation_mother + education_self:education_father +
  education_self:education_mother + occupation_self:occupation_father +
  occupation_self:occupation_mother + income_hh_eq + financial_difficulties +
  hh_size + children_yn

# Model 5f: Interaction Between Parental SEP Variables


model_5f <- dhd_index ~ education_self + occupation_self +
  education_father * occupation_father + education_mother * occupation_mother +
  education_self:education_father + education_self:education_mother +
  occupation_self:occupation_father + occupation_self:occupation_mother +
  income_hh_eq + financial_difficulties

# Model 5g: Add Home Ownership Status


model_5g <- dhd_index ~ education_self + occupation_self +
  education_father + occupation_father + education_mother +
  occupation_mother + education_self:education_father +
  education_self:education_mother + occupation_self:occupation_father +
  occupation_self:occupation_mother + income_hh_eq + financial_difficulties +
  home_owner

# Model 5h: Include Diet-Related Variables


model_5h <- dhd_index ~ education_self + occupation_self +
  education_father + occupation_father + education_mother +
  occupation_mother + education_self:education_father +
  education_self:education_mother + occupation_self:occupation_father +
  occupation_self:occupation_mother + income_hh_eq + financial_difficulties +
  dhd_kcal

# Model 5i: Interaction Between Financial Difficulties and Parental SEP


model_5i <- dhd_index ~ education_self + occupation_self +
  education_father + occupation_father + education_mother +
  occupation_mother + education_self:education_father +
  education_self:education_mother + occupation_self:occupation_father +
  occupation_self:occupation_mother + income_hh_eq + financial_difficulties +
  financial_difficulties:education_father +
  financial_difficulties:education_mother

# Model 5j: Interaction Between All SEP Variables


model_5j <- dhd_index ~ education_self * occupation_self *
  education_father * occupation_father *
  education_mother * occupation_mother +
  income_hh_eq + financial_difficulties

# Model 6: Gender as a Moderator Between SEP and Diet Quality

# Model 6a: Interaction Between Income and Gender


model_6a <- dhd_index ~ education_self + occupation_self + woman +
  education_father + occupation_father + education_mother + occupation_mother +
  income_hh_eq + financial_difficulties + income_hh_eq:woman

# Model 6b: Interaction Between Financial Difficulties and Gender


model_6b <- dhd_index ~ education_self + occupation_self + woman +
  education_father + occupation_father + education_mother + occupation_mother +
  income_hh_eq + financial_difficulties + financial_difficulties:woman

# Model 6c: Include Age and Smoking Status


model_6c <- dhd_index ~ education_self + occupation_self + woman +
  education_father + occupation_father + education_mother + occupation_mother +
  income_hh_eq + financial_difficulties + age + smoking_status

# Model 6d: Interaction Between Parental SEP and Gender Only


model_6d <- dhd_index ~ education_self + occupation_self + woman +
  education_father:woman + occupation_father:woman +
  education_mother:woman + occupation_mother:woman +
  income_hh_eq + financial_difficulties

# Model 6e: Include Household Size and Children


model_6e <- dhd_index ~ education_self + occupation_self + woman +
  education_father + occupation_father + education_mother + occupation_mother +
  income_hh_eq + financial_difficulties + hh_size + children_yn

# Model 6f: Interaction Between Individual SEP and Gender Only


model_6f <- dhd_index ~ education_self:woman + occupation_self:woman +
  education_father + occupation_father + education_mother + occupation_mother +
  income_hh_eq + financial_difficulties

# Model 6g: Add Diet-Related Variables


model_6g <- dhd_index ~ education_self + occupation_self + woman +
  education_father + occupation_father + education_mother + occupation_mother +
  income_hh_eq + financial_difficulties + dhd_kcal

# Model 6h: Interaction Between All SEP Variables and Gender


model_6h <- dhd_index ~ (education_self + occupation_self +
  education_father + occupation_father + education_mother + occupation_mother) *
  woman + income_hh_eq + financial_difficulties

# Model 6i: Interaction Between Education and Gender


model_6i <- dhd_index ~ education_self * woman + occupation_self +
  education_father * woman + education_mother * woman +
  income_hh_eq + financial_difficulties

# Model 6j: Include Home Ownership Status


model_6j <- dhd_index ~ education_self + occupation_self + woman +
  education_father + occupation_father + education_mother + occupation_mother +
  income_hh_eq + financial_difficulties + home_owner

# Model 7: Interactions Between Individual and Partner’s SEP

# Model 7a: Add Interaction Between Partner’s Education and Gender


model_7a <- dhd_index ~ education_self + occupation_self +
  education_partner + occupation_partner + woman +
  education_partner:woman + income_hh_eq + financial_difficulties

# Model 7b: Interaction Between Partner’s Occupation and Gender


model_7b <- dhd_index ~ education_self + occupation_self +
  education_partner + occupation_partner + woman +
  occupation_partner:woman + income_hh_eq + financial_difficulties

# Model 7c: Include Age and Smoking Status


model_7c <- dhd_index ~ education_self + occupation_self +
  education_partner + occupation_partner + income_hh_eq +
  financial_difficulties + age + smoking_status

# Model 7d: Interaction Between Education and Occupation (Self and Partner)


model_7d <- dhd_index ~ education_self * occupation_self +
  education_partner * occupation_partner +
  education_self:education_partner + occupation_self:occupation_partner +
  income_hh_eq + financial_difficulties

# Model 7e: Include Household Size and Children


model_7e <- dhd_index ~ education_self + occupation_self +
  education_partner + occupation_partner +
  education_self:education_partner + occupation_self:occupation_partner +
  income_hh_eq + financial_difficulties + hh_size + children_yn

# Model 7f: Interaction Between Partner’s SEP Variables


model_7f <- dhd_index ~ education_self + occupation_self +
  education_partner * occupation_partner +
  education_self:education_partner + occupation_self:occupation_partner +
  income_hh_eq + financial_difficulties

# Model 7g: Add Home Ownership Status


model_7g <- dhd_index ~ education_self + occupation_self +
  education_partner + occupation_partner +
  education_self:education_partner + occupation_self:occupation_partner +
  income_hh_eq + financial_difficulties + home_owner

# Model 7h: Include Diet-Related Variables


model_7h <- dhd_index ~ education_self + occupation_self +
  education_partner + occupation_partner +
  education_self:education_partner + occupation_self:occupation_partner +
  income_hh_eq + financial_difficulties + dhd_kcal

# Model 7i: Interaction Between Financial Difficulties and Partner’s SEP


model_7i <- dhd_index ~ education_self + occupation_self +
  education_partner + occupation_partner +
  education_self:education_partner + occupation_self:occupation_partner +
  income_hh_eq + financial_difficulties +
  financial_difficulties:education_partner +
  financial_difficulties:occupation_partner

# Model 7j: Interaction Between All SEP Variables


model_7j <- dhd_index ~ education_self * occupation_self *
  education_partner * occupation_partner +
  income_hh_eq + financial_difficulties

# Model 8: Interactions with Partner’s SEP and Gender

# Model 8a: Interaction Between Partner’s SEP and Gender Only


model_8a <- dhd_index ~ education_self + occupation_self + woman +
  education_partner:woman + occupation_partner:woman +
  income_hh_eq + financial_difficulties

# Model 8b: Include Age and Smoking Status


model_8b <- dhd_index ~ education_self + occupation_self + woman +
  education_partner + occupation_partner +
  education_self:education_partner + occupation_self:occupation_partner +
  education_self:woman + occupation_self:woman +
  education_partner:woman + occupation_partner:woman +
  income_hh_eq + financial_difficulties + age + smoking_status

# Model 8c: Interaction Between Income and Gender


model_8c <- dhd_index ~ education_self + occupation_self + woman +
  education_partner + occupation_partner +
  income_hh_eq * woman + financial_difficulties

# Model 8d: Include Household Size and Children


model_8d <- dhd_index ~ education_self + occupation_self + woman +
  education_partner + occupation_partner +
  education_self:education_partner + occupation_self:occupation_partner +
  education_self:woman + occupation_self:woman +
  education_partner:woman + occupation_partner:woman +
  income_hh_eq + financial_difficulties + hh_size + children_yn

# Model 8e: Interaction Between Financial Difficulties and Gender


model_8e <- dhd_index ~ education_self + occupation_self + woman +
  education_partner + occupation_partner +
  financial_difficulties * woman +
  education_self:woman + occupation_self:woman +
  education_partner:woman + occupation_partner:woman +
  income_hh_eq

# Model 8f: Interaction Between All SEP Variables and Gender


model_8f <- dhd_index ~ (education_self + occupation_self +
  education_partner + occupation_partner) * woman +
  education_self:education_partner + occupation_self:occupation_partner +
  income_hh_eq + financial_difficulties

# Model 8g: Add Home Ownership Status


model_8g <- dhd_index ~ education_self + occupation_self + woman +
  education_partner + occupation_partner +
  education_self:education_partner + occupation_self:occupation_partner +
  education_self:woman + occupation_self:woman +
  education_partner:woman + occupation_partner:woman +
  income_hh_eq + financial_difficulties + home_owner

# Model 8h: Include Diet-Related Variables


model_8h <- dhd_index ~ education_self + occupation_self + woman +
  education_partner + occupation_partner +
  education_self:education_partner + occupation_self:occupation_partner +
  education_self:woman + occupation_self:woman +
  education_partner:woman + occupation_partner:woman +
  income_hh_eq + financial_difficulties + dhd_kcal

# Model 8i: Interaction Between Partner’s SEP and Financial Difficulties


model_8i <- dhd_index ~ education_self + occupation_self + woman +
  education_partner + occupation_partner +
  financial_difficulties:education_partner +
  financial_difficulties:occupation_partner +
  education_self:woman + occupation_self:woman +
  education_partner:woman + occupation_partner:woman +
  income_hh_eq + financial_difficulties

# Model 8j: Interaction Between All Variables


model_8j <- dhd_index ~ (education_self * occupation_self *
  education_partner * occupation_partner * woman) +
  income_hh_eq + financial_difficulties

# Model 9: Comprehensive Multivariate # Model with Additional Covariates and Interactions

# Model 9a: Add Interaction Between Age and Gender


model_9a <- dhd_index ~ education_self + occupation_self + woman +
  age * woman + smoking_status + dhd_kcal + hh_size + children_yn +
  education_father + occupation_father + education_mother + occupation_mother +
  education_partner + occupation_partner +
  education_self:education_father + education_self:education_mother +
  occupation_self:occupation_father + occupation_self:occupation_mother +
  education_self:education_partner + occupation_self:occupation_partner +
  education_self:woman + occupation_self:woman +
  income_hh_eq + financial_difficulties

# Model 9b: Include Interaction Between Smoking Status and Gender


model_9b <- dhd_index ~ education_self + occupation_self + woman +
  age + smoking_status * woman + dhd_kcal + hh_size + children_yn +
  education_father + occupation_father + education_mother + occupation_mother +
  education_partner + occupation_partner +
  education_self:education_father + education_self:education_mother +
  occupation_self:occupation_father + occupation_self:occupation_mother +
  education_self:education_partner + occupation_self:occupation_partner +
  education_self:woman + occupation_self:woman +
  income_hh_eq + financial_difficulties

# Model 9c: Add Home Ownership Status


model_9c <- dhd_index ~ education_self + occupation_self + woman +
  age + smoking_status + dhd_kcal + hh_size + children_yn + home_owner +
  education_father + occupation_father + education_mother + occupation_mother +
  education_partner + occupation_partner +
  education_self:education_father + education_self:education_mother +
  occupation_self:occupation_father + occupation_self:occupation_mother +
  education_self:education_partner + occupation_self:occupation_partner +
  education_self:woman + occupation_self:woman +
  income_hh_eq + financial_difficulties

# Model 9d: Interaction Between Financial Difficulties and Gender


model_9d <- dhd_index ~ education_self + occupation_self + woman +
  age + smoking_status + dhd_kcal + hh_size + children_yn +
  education_father + occupation_father + education_mother + occupation_mother +
  education_partner + occupation_partner +
  education_self:education_father + education_self:education_mother +
  occupation_self:occupation_father + occupation_self:occupation_mother +
  education_self:education_partner + occupation_self:occupation_partner +
  education_self:woman + occupation_self:woman +
  financial_difficulties * woman + income_hh_eq

# Model 9e: Include Interaction Between Income and Gender


model_9e <- dhd_index ~ education_self + occupation_self + woman +
  age + smoking_status + dhd_kcal + hh_size + children_yn +
  education_father + occupation_father + education_mother + occupation_mother +
  education_partner + occupation_partner +
  education_self:education_father + education_self:education_mother +
  occupation_self:occupation_father + occupation_self:occupation_mother +
  education_self:education_partner + occupation_self:occupation_partner +
  education_self:woman + occupation_self:woman +
  income_hh_eq * woman + financial_difficulties

# Model 9f: Interaction Between All SEP Variables and Gender


model_9f <- dhd_index ~ (education_self + occupation_self +
  education_father + occupation_father + education_mother + occupation_mother +
  education_partner + occupation_partner) * woman +
  age + smoking_status + dhd_kcal + hh_size + children_yn +
  education_self:education_father + education_self:education_mother +
  occupation_self:occupation_father + occupation_self:occupation_mother +
  education_self:education_partner + occupation_self:occupation_partner +
  income_hh_eq + financial_difficulties

# Model 9g: Add Diet Quality Index Components


model_9g <- dhd_index ~ education_self + occupation_self + woman +
  age + smoking_status + hh_size + children_yn +
  dhd_kcal + dhd_sugar + dhd_fruit + dhd_vegetables +
  education_father + occupation_father + education_mother + occupation_mother +
  education_partner + occupation_partner +
  education_self:education_father + education_self:education_mother +
  occupation_self:occupation_father + occupation_self:occupation_mother +
  education_self:education_partner + occupation_self:occupation_partner +
  education_self:woman + occupation_self:woman +
  income_hh_eq + financial_difficulties

# Model 9h: Include Interaction Between Household Size and Children


model_9h <- dhd_index ~ education_self + occupation_self + woman +
  age + smoking_status + dhd_kcal + hh_size * children_yn +
  education_father + occupation_father + education_mother + occupation_mother +
  education_partner + occupation_partner +
  education_self:education_father + education_self:education_mother +
  occupation_self:occupation_father + occupation_self:occupation_mother +
  education_self:education_partner + occupation_self:occupation_partner +
  education_self:woman + occupation_self:woman +
  income_hh_eq + financial_difficulties

# Model 9i: Interaction Between Age and Smoking Status


model_9i <- dhd_index ~ education_self + occupation_self + woman +
  age * smoking_status + dhd_kcal + hh_size + children_yn +
  education_father + occupation_father + education_mother + occupation_mother +
  education_partner + occupation_partner +
  education_self:education_father + education_self:education_mother +
  occupation_self:occupation_father + occupation_self:occupation_mother +
  education_self:education_partner + occupation_self:occupation_partner +
  education_self:woman + occupation_self:woman +
  income_hh_eq + financial_difficulties

# Model 9j: Comprehensive # Model with All Interactions


model_9j <- dhd_index ~ (education_self * occupation_self *
  education_father * occupation_father *
  education_mother * occupation_mother *
  education_partner * occupation_partner * woman) +
  age + smoking_status + dhd_kcal + hh_size + children_yn +
  income_hh_eq + financial_difficulties

# paste0("model_", rep(1:9, each = 10), letters[1:10]) %>% toString()
```
```{r}
# Create a named list of models
# model_list <- named_list(
#   model_1, model_2, model_3,
#   model_4, model_5, model_6,
#   model_7, model_8, model_9
# )

model_list <- named_list(
  model_1a, model_1b, model_1c, model_1d, model_1e, model_1f, model_1g, model_1h, model_1i, model_1j, model_2a, model_2b, model_2c, model_2d, model_2e, model_2f, model_2g, model_2h, model_2i, model_2j, model_3a, model_3b, model_3c, model_3d, model_3e, model_3f, model_3g, model_3h, model_3i, model_3j, model_4a, model_4b, model_4c, model_4d, model_4e, model_4f, model_4g, model_4h, model_4i, model_4j, model_5a, model_5b, model_5c, model_5d, model_5e, model_5f, model_5g, model_5h, model_5i, model_5j, model_6a, model_6b, model_6c, model_6d, model_6e, model_6f, model_6g, model_6h, model_6i, model_6j, model_7a, model_7b, model_7c, model_7d, model_7e, model_7f, model_7g, model_7h, model_7i, model_7j, model_8a, model_8b, model_8c, model_8d, model_8e, model_8f, model_8g, model_8h, model_8i, model_8j, model_9a, model_9b, model_9c, model_9d, model_9e, model_9f, model_9g, model_9h, model_9i, model_9j
)

# Function to extract numeric data based on the variables in model_9
get_numeric_data <- function(df, model = model_9) {
  df[as.character(attr(terms(model), "variables"))[-1]]
}

data <- unadjusted #get_numeric_data(adjusted %||% unadjusted)
```

```{r}
#| label: prepare-parallel-computing
#| echo: FALSE

if (is.null(adjusted) || is.null(unadjusted)) {
  num_cores <- parallel::detectCores() - 1
  cluster <- parallel::makeCluster(num_cores)
  invisible(parallel::clusterEvalQ(cluster, {
    library(boot, quietly = TRUE)
    library(robustbase, quietly = TRUE)
    library(dplyr, quietly = TRUE)
  }))

  parallel::clusterExport(
    cluster,
    varlist = c("calculate_p_value_from_ci"),
    envir = environment()
  )

  t1 <- Sys.time()
  fits <- list()
  for (model_name in names(model_list)) {
    model_start_time <- Sys.time()
    cumulative_time <- difftime(model_start_time, t1, units = "mins")
    cat(paste0(
      "Fitting model ", dQuote(model_name),
      " (elapsed time: ", round(cumulative_time, 2), " mins)\n"
    ))

    model_formula <- model_list[[model_name]]
    fits[[model_name]] <- tryCatch(
      {
        bootstrap(
          data = data,
          formula = model_formula,
          seed = 123,
          num_bootstraps = N_BOOTSTRAPS,
          cluster = cluster
        )
      },
      error = function(e) {
        cli::cli_alert_warning(paste0("Error in model ", dQuote(model_name), ": ", e$message, "\n"))
        list()
      }
    )


    parallel::stopCluster(cluster)
    save_rds(fits,
      name = paste0(ifelse(is.null(adjusted), "unadjusted_alt", "adjusted_detailed_alt"), "_fits"),
      dir = "data"
    )
  }
}
```

After the unadjusted model fits, we evaluate the assumptions to identify high leverage points. Subsequently, we bootstrap with the adjusted data to re-fit the models.



```{r}
tryCatch(
  {
    closeAllConnections()
  },
  error = function(e) {

  }
)

data <- adjusted # get_numeric_data(adjusted)

num_cores <- parallel::detectCores() - 1
cluster <- parallel::makeCluster(num_cores)
invisible(parallel::clusterEvalQ(cluster, {
  library(boot, quietly = TRUE)
  library(robustbase, quietly = TRUE)
  library(dplyr, quietly = TRUE)
}))

parallel::clusterExport(cluster,
  varlist = c("calculate_p_value_from_ci"),
  envir = environment()
)

t1 <- Sys.time()
fits <- list()
for (model_name in names(model_list)) {
  model_start_time <- Sys.time()
  cumulative_time <- difftime(model_start_time, t1, units = "mins")
  cat(paste0(
    "Fitting model ",
    dQuote(model_name),
    " (elapsed time: ",
    round(cumulative_time, 2),
    " mins)\n"
  ))

  fits[[model_name]] <- tryCatch(
    {
      bootstrap(
        data = data,
        formula = model_formula,
        seed = 123,
        num_bootstraps = N_BOOTSTRAPS,
        cluster = cluster
      )
    },
    error = function(e) {
      cli::cli_alert_warning(paste0("Error in model ", dQuote(model_name), ": ", e$message, "\n"))
      list()
    }
  )
}

parallel::stopCluster(cluster)

save_rds(fits, name = "adjusted_alt_fits", dir = "data")

```



```{r}
#| label: structure-models
#| echo: FALSE
#| message: FALSE
model_fits <- structure_models()
```

## Sensitivity Analysis: Comparing Robust Linear Regression Models with and without Influential Observations

Identifying influential observations through regression diagnostics, such as Cook's distance, in conjunction with robust estimation methods, such as the MM-estimator, is essential for ensuring the validity and reliability of results in robust linear regression models. A comprehensive analysis of the 13 robust linear regression models using the MM estimator and Cook's distance, a measure of influence, identified seven data points (observations: 361, 374, 540, 977, 1042, 1175, and 1217) that consistently exceeded the typical threshold of $\frac{4}{n}$ (where $n$ is the sample size) in at least four distinct model specifications. 
Univariate and multivariate outlier analyses conducted before modeling (see p.4) was unable to detect these influential observations due to their unique configuration of predictor and response values, as an observation's influence in a robust regression model is determined by the combination of its leverage (the extremity of its predictor values) and discrepancy (the degree to which the model poorly fits the observation) rather than its univariate or multivariate distance from the center of the data.
To quantify the impact of these influential points, a sensitivity analysis was performed by fitting the robust MM-estimator models with and without the seven observations. The results revealed that the inclusion or exclusion of these data points had a substantial effect on key model parameters and goodness-of-fit measures, such as the coefficient estimates ($\hat{\beta}$) and adjusted R-squared ($\bar{R}^2$), even when using a robust estimation method designed to mitigate the influence of outliers. This finding provides direct evidence that these observations drive the results considerably, potentially obscuring the proper relationships between the predictors and the response variable. The recurrent high influence of these seven observations across models with different predictors underscores that their impact stems from their overall configuration of predictor and response values rather than a specific relationship with a single predictor.
Although the evidence supports the removal of these seven observations, it is crucial to acknowledge the potential limitations and biases introduced by data exclusion, even in the context of robust linear regression models. Therefore, the results are reported with and without these observations to maintain transparency and provide a comprehensive understanding of the data. By prioritizing regression diagnostics in conjunction with robust estimation techniques, analysts can identify truly influential observations and make informed choices to ensure the accuracy and validity of their results. This approach enables a more nuanced understanding of the data. It helps prevent erroneous conclusions based on the undue influence of a small subset of observations, ultimately strengthening the reliability and interpretability of robust linear regression models.


```{r}
#| label: fits-performance-calculation
#| message: false
#| warning: false
#| include: false
fits_df <- calculate_performance(model_fits)
move_column_after(fits_df, "r2_p_value", "R2") %>%
  dplyr::mutate(r2_p_value = format_p_value(
    r2_p_value,
    add_stars = TRUE, omit_p_label = FALSE
  )) %>%
  dplyr::arrange(model, dplyr::desc(method))
fits_df
```


```{r}
#| label: model-fits-table
#| tbl-cap: "Comparison of Robust Linear Regression Models with and without Leverage Points"
col_names <- c(
  "Leverage", "Method", "$R^2$", "", "$R^2 (adj.)$", "RMSE", "$\\sigma$", "AIC", "AICc", "BIC"
)
footnote_text <- paste(
  "The ($R^2$) and adjusted ($R^2$ adj.) gauge model fit,",
  "RMSE evaluates residual magnitude, and $\\sigma$ denotes error standard deviation.",
  "The Akaike (AIC) and Bayesian (BIC) information criteria balance fit and parsimony (lower is better),",
  "but apply only to OLS models, not robust models, as the latter utilize M-estimation without assuming normal residuals.",
  "AICc adjusts AIC for small sample sizes."
)

fits_df %>%
  remove_columns_by_name("model") %>%
  as_kable() %>%
  kableExtra::pack_rows(
    index = c(
      "Model 1" = 4,
      "Model 2" = 4,
      "Model 3" = 4,
      "Model 4" = 4,
      "Model 5" = 4,
      "Model 6" = 4,
      "Model 7" = 4,
      "Model 8" = 4,
      "Model 9" = 4
    )
  ) %>%
  footnote() %>%
  add_style()
```

```{r include=FALSE}
nested_caption <- paste(
  "Wald and",
  c("Deviance", "Explained Variance"),
  "Tests for Nested",
  c("Robust", "OLS"),
  "Linear Regression Models (Models 1-9, adjusted for influential observations)"
)

lmrob_column_headers <- c(
  "Reference Model",
  "$DF_{\\text{pseudo1}}$",
  "$DF_{\\text{pseudo2}}$",
  "Wald",
  " ",
  "Deviance",
  " ",
  "DF"
)

ols_column_headers <- c(
  "Reference Model",
  "Res. $DF_1$",
  "Res. $DF_2$",
  "Wald",
  " ",
  "$R^2 \\Delta$",
  " ",
  "DF"
)

lmrob_footnote <- paste(
  "$^{***} p \\leq 0.001;\\ ^{**} p \\leq 0.01;\\ ^{*} p \\leq 0.05$.",
  "The pseudo-degrees of freedom (DF$_{\\text{pseudo}}$) represent an adjusted",
  "count of degrees of freedom in each robust model. The Wald and Deviance test",
  "statistics measure the improvement of the full model over the restricted model.",
  "The degrees of freedom (DF) represent the difference in the number of",
  "parameters between models. The p-values indicate the probability of observing",
  "test statistics as extreme as those computed under the null hypothesis,",
  "which assumes equivalence between the two models. Non-nested pairs",
  "(`—`) are included for completeness."
)

ols_footnote <- paste(
  "Residual degrees of freedom (Res. DF) represent the difference in the number",
  "of parameters between the full and restricted models. The Wald test statistic",
  "measures the improvement of the full model over the restricted model.",
  "The $R^2$ difference ($R^2 \\Delta$) quantifies the change in model fit.",
  "The degrees of freedom (DF) represent the difference in the number of parameters.",
  "The p-values indicate the probability of observing test statistics as extreme",
  "as those computed under the null hypothesis, which assumes equivalence",
  "between the two models. Non-nested pairs (`—`) are included for completeness."
)
```

```{r}
#| label: nested-models-comparison
#| tbl-cap:  "Wald and Deviance Tests for Nested Robust Linear Regression Models (Models 1-9, adjusted for influential observations)"
nested_models_comparison <- compare_nested_models(model_fits, "adjusted", "robust")
model_counts <- as.numeric(table(nested_models_comparison$A))

nested_models_comparison %>%
  remove_columns_by_name("A", "Models") %>%
  as_kable(
    align = "rrrrlrll",
    cols = lmrob_column_headers
  ) %>%
  kableExtra::pack_rows(
    index = c(
      "Model 1" = model_counts[1],
      "Model 2" = model_counts[2],
      "Model 3" = model_counts[3],
      "Model 4" = model_counts[4],
      "Model 5" = model_counts[5],
      "Model 6" = model_counts[6],
      "Model 7" = model_counts[7],
      "Model 8" = model_counts[8]
    )
  ) %>%
footnote(lmrob_footnote)
```

```{r}
#| label: nested-models-comparison-ols
#| tbl-cap: "Wald and Explained Variance Tests for Nested OLS Linear Regression Models (Models 1-9, adjusted for influential observations)"
nested_models_comparison_ols <- compare_nested_models(model_fits, "adjusted", "ols")
model_counts_ols <- as.numeric(table(nested_models_comparison_ols$A))

nested_models_comparison_ols %>%
  remove_columns_by_name("A", "Models") %>%
  as_kable(
    align = "rrrrlrll",
    cols = ols_column_headers
  ) %>%
  kableExtra::pack_rows(
    index = c(
      "Model 1" = model_counts_ols[1],
      "Model 2" = model_counts_ols[2],
      "Model 3" = model_counts_ols[3],
      "Model 4" = model_counts_ols[4],
      "Model 5" = model_counts_ols[5],
      "Model 6" = model_counts_ols[6],
      "Model 7" = model_counts_ols[7],
      "Model 8" = model_counts_ols[8]
    )
  ) %>%
  footnote(ols_footnote)
```

```{r simplify_model_fits, include=FALSE}
fits <- purrr::map(model_fits$adjusted$robust, ~ .x$fit)
```

```{r jn_plots_model_5, fig.cap="Johnson-Neyman Plots (Model 3)", message=FALSE, warning=FALSE}
jn_plot(fits$model_3) %>%
  Filter(f = length) %>%
  gridExtra::grid.arrange(
    grobs = .,
    bottom = "Model 3 (robust regression without influential observations)",
    ncol = 2
  )
```

```{r jn_plots_model_5, fig.cap="Johnson-Neyman Plots (Model 5)", message=FALSE, warning=FALSE}
jn_plot(fits$model_5) %>%
  Filter(f = length) %>%
  gridExtra::grid.arrange(
    grobs = .,
    bottom = "Model 5 (robust regression without influential observations)",
    ncol = 2
  )
```

```{r jn_plots_model_6, fig.cap="Johnson-Neyman Plots (Model 6)", message=FALSE, warning=FALSE}
jn_plot(fits$model_6) %>%
  Filter(f = length) %>%
  gridExtra::grid.arrange(
    grobs = .,
    bottom = "Model 6 (robust regression without influential observations)",
    ncol = 2
  )
```

```{r jn_plots_model_7, fig.cap="Johnson-Neyman Plots (Model 7)", message=FALSE, warning=FALSE}
jn_plot(fits$model_7) %>%
  Filter(f = length) %>%
  gridExtra::grid.arrange(
    grobs = .,
    bottom = "Model 7 (robust regression without influential observations)",
    ncol = 2
  )
```

```{r jn_plots_model_8, fig.cap="Johnson-Neyman Plots (Model 8)", message=FALSE, warning=FALSE}
jn_plot(fits$model_8) %>%
  Filter(f = length) %>%
  gridExtra::grid.arrange(
    grobs = .,
    bottom = "Model 8 (robust regression without influential observations)",
    ncol = 2
  )
```

```{r jn_plots_model_9, fig.cap="Johnson-Neyman Plots (Model 9)", message=FALSE, warning=FALSE}
jn_plot(fits$model_9) %>%
  Filter(f = length) %>%
  gridExtra::grid.arrange(
    grobs = .,
    bottom = "Model 9 (robust regression without influential observations)",
    ncol = 2
  )
```

```{r adjusted_robust_model_summary}
model_summary(model_fits$adjusted$robust, adjusted = TRUE)
model_summary(model_fits$adjusted$ols, adjusted = TRUE)
```

```{r unadjusted_robust_model_summary}
model_summary(model_fits$unadjusted$robust, adjusted = FALSE)
model_summary(model_fits$unadjusted$ols, adjusted = FALSE)
```



# Appendix

Full tables for the robust linear regression models with and without influential observations are provided below. Each table includes the coefficient estimates, standard errors, t-values, p-values, and 95% confidence intervals for each predictor in the model. The tables also present the bootstrapped coefficient estimates, standard errors, bias, p-values, and 95% confidence intervals for each predictor. The bootstrapped results provide additional information on the stability and reliability of the coefficient estimates in the models.

```{r}
# contains coef, se, t, p, ci_lower, ci_upper, boot_coef, boot_se, boot_bias, boot_p, boot_ci_lower, boot_ci_upper
```
